Modular properties of string, branching and dual functions

### 2010-07-18 AffineLieAlgebras/DualFunctions/ModularProperties.tex
Some work

\begin{abstract}
  We review the appearance of modular transformation in the representation theory of affine Lie algebras, especially the modular properties of branching and string function.
\end{abstract}
\begin{Def}
  {\bf Modular transformation}
  \begin{equation}
    \label{eq:1}
    A=
    \begin{pmatrix} a & b\\ c & d
    \end{pmatrix} \quad\mbox{where}\; a,b,c,d\in\mathbb{Z},\quad ad-bc=1,
  \end{equation}
  \begin{equation}
    \label{eq:2}
    z\to A z=\frac{az+b}{cz+d}
  \end{equation}
\end{Def}
\begin{Def}
  {\bf Modular form} of weight $k$ is the analytic function $F$ on the upper half plane with infinity ( $F(i\infty)$ is holomorphic) which transforms in the following way under the modular transformations
  \begin{equation}
    \label{eq:3}
    F(Az)=(cz+d)^k F(z)
  \end{equation}
\end{Def}

@@

For $k\geq 4$ we have the following example (Eisenstein series):
\begin{equation}
  \label{eq:5}
  G_k(\Lambda)=\sum_{\lambda\in \Lambda, \lambda\neq 0} \lambda^{-k}
\end{equation}
Vector space of weight $k$ modular forms is finite dimensional.

The quotient of the complex numbers by a lattice is an elliptic curve and every elliptic curve arises in this way. ($y^2=f(x)$, $f$ - cubic polynomial).

For the $d$-dimensional lattice $L$ with the non-degenerate form $(\cdot,\cdot)$ we can define theta-functions
\begin{equation}
  \label{eq:8}
  \Theta_L (\tau)=\sum_{x\in L}e^{\pi i \tau (x,x)}
\end{equation}
$\Theta_L$ is modular of weight $d/2$.

@@

Now we can consider the weight lattice of the affine Lie algebra $\mathfrak{g}$.
Cartan subalgebra:
\begin{equation}
  \label{eq:6}
  \mathfrak{h}=\mathbb{C}\Lambda_0\oplus \co{\mathfrak{h}}\oplus \mathbb{C}\delta
\end{equation}
$h\in \mathfrak{h}$
\begin{equation}
  \label{eq:7}
  h=(z;\tau;t)=2\pi i (-\tau \Lambda_0 +z + t\delta)\quad (\tau,t\in \mathbb{Z}, z\in \co{\mathfrak{h}})
\end{equation}

@@

See Kac, Peterson 1984 for the original proof.
\section{Questions}
\label{sec:questions}
\begin{itemize}
\item The modular properties of WZW models are naturally connected with the problem of theory definition on the torus. What can we say about the pure representation theory? What is the analogue of the torus?
\item What kind of constrains on the stars and fans follow from the modular invariance?
\item The modular invariance is important in the number theory. What problems of number theory are connected through the modular transformations with the representation theory?
\item I need to write about the Hecke algebras
\item How are Hecke algebras connected to the modular forms?
\item Are there any kind of modular invariance analogues for the finite-dimensional Lie algebras? There exists the lattice.
\item What become of modular invariance during the quantization? Are there any properties of quantum integrable systems that follow from the modular invariance?
\end{itemize}

>>>>

\begin{abstract}
  We review the appearance of modular transformation in the representation theory of affine Lie algebras, especially the modular properties of branching and string function.
\end{abstract}
\begin{Def}
  {\bf Modular transformation}
  \begin{equation}
    \label{eq:1}
    A=
    \begin{pmatrix} a & b\\ c & d
    \end{pmatrix} \quad\mbox{where}\; a,b,c,d\in\mathbb{Z},\quad ad-bc=1,
  \end{equation}
  \begin{equation}
    \label{eq:2}
    z\to A z=\frac{az+b}{cz+d}
  \end{equation}
\end{Def}
\begin{Def}
  {\bf Modular form} of weight $k$ is the function $F$ on the upper half plane with infinity which transforms in the following way under the modular transformations
  \begin{equation}
    \label{eq:3}
    F(Az)=(cz+d)^k F(z)
  \end{equation}
\end{Def}

@@

For $k\geq 4$ we have the following example (Eisenstein series):
\begin{equation}
  \label{eq:5}
  G_k(\Lambda)=\sum_{\lambda\in \Lambda, \lambda\neq 0} \lambda^{-k}
\end{equation}
\begin{comment}
Eisenstein series are important in the discussion of $\Gamma_n$-modular forms, see later.
\end{comment}

This requirement is not really strong, additional requirement of $F$ being holomorphic function at the upper half-plane with infinity is really powerful. It is well-known that the space of the (holomorphic) modular forms of weight $k$ is finite-dimensional. We discuss it later in the connection with the cusp forms and forms of weight $k$ over $\Gamma(N)$.

\begin{comment}
The quotient of the complex numbers by a lattice is an elliptic curve and every elliptic curve arises in this way. ($y^2=f(x)$, $f$ - cubic polynomial).  
\end{comment}

\begin{Def}
  {\bf Index} of a subgroup $H\subset G$ is the number of cosets of $H$ in $G$, it is denoted by $\left|G:H\right|$.
\end{Def}.

Let $\Gamma \subset \mathrm{SL}(2,\mathbb{Z})$ be a subgroup of finite index. Given a positive integer $k$, a modular form $F$ for $\Gamma$ of weight $k$ is a holomorphic function on the upper half-plane which transforms as in \eqref{eq:3} and is also holomorphic at the cusps of $\Gamma$.

For $N\in \mathbb{N},$ define
\begin{equation}
  \label{eq:12}
  \Gamma(N)=\left\{
      \begin{pmatrix} a & b \\ c & d
      \end{pmatrix}
      \equiv
      \begin{pmatrix}
        1 & 0 \\
        0 & 1
      \end{pmatrix} (\text{mod}\; N)
      \right|\left.
        \begin{matrix}
          a,b,c,d\in \mathbb{Z},\quad ad-bc=1\\
          a\equiv d\equiv 1\; \text{mod}\; N,\quad b\equiv c\equiv 0\; \text{mod}\; N
        \end{matrix}
      \right\}
  \end{equation}
The subgroup $\Gamma(N)$ has {\it finite index} in $\Gamma(1)=SL(2,\mathbb{Z})$. For any $\tau\in \mathbb{C}$ there exists element of $SL(2,\mathbb{Z})$, which transforms it to the fundamental domain.

For $A\in SL(2,\mathbb{Z})$ define
\begin{equation}
  \label{eq:13}
  j(A,\tau)=c\tau+d.
\end{equation}
Note that
\begin{equation}
  \label{eq:14}
  j(BA,\tau)=j(B,A\tau)j(A,\tau).
\end{equation}
For any subgroup $\Gamma$ of finite index define a {\bf multiplier system} $\chi$ as a mapping
\begin{equation}
  \label{eq:15}
  \chi:\Gamma\to U(1).
\end{equation}
Let $k\in \mathbb{R}$. A complex valued holomorphic function $f$ on the upper half-plane of the complex plane is a modular form of weight $k$ with respect to $\Gamma$ if there exists a multiplier system $\chi$ for $\Gamma$ with
\begin{equation}
  \label{eq:16}
  f(\tau)=\chi(A) j(A,\tau)^{-k} f(A\tau),
\end{equation}
for all $\tau$ in the upper half-plane and $A\in \Gamma$.

For the $d$-dimensional lattice $L$ with the non-degenerate form $(\cdot,\cdot)$ we can define theta functions
\begin{equation}
  \label{eq:8}
  \Theta_L (\tau)=\sum_{x\in L}e^{\pi i \tau (x,x)}
\end{equation}
$\Theta_L$ is modular of weight $d/2$.

@@

Since this construction gives us the modular forms from any lattice it is natural to introduce theta functions on the weight lattice of affine Lie algebra.

The connection between the theta functions and the string/branching functions is revealed in the papers \cite{kac1984infinite,kac1988modular}. Here we briefly describe this connection using the text of the book \cite{kass1990ala} (see pages 212-218).

Now we consider the weight lattice of the affine Lie algebra $\mathfrak{g}$.
Its Cartan subalgebra has the following decomposition:
\begin{equation}
  \label{eq:6}
  \mathfrak{h}=\mathbb{C}\Lambda_0\oplus \co{\mathfrak{h}}\oplus \mathbb{C}\delta
\end{equation}
$h\in \mathfrak{h}$
\begin{equation}
  \label{eq:7}
  h=(z;\tau;t)=2\pi i (-\tau \Lambda_0 +z + t\delta)\quad (\tau,t\in \mathbb{Z}, z\in \co{\mathfrak{h}})
\end{equation}

@@

\section{Questions}
\label{sec:questions}
\begin{itemize}
\item The modular properties of WZW models are naturally connected with the problem of theory definition on the torus. What can we say about the pure representation theory? What is the analogue of the torus?
\item What kind of constrains on the stars and fans follow from the modular invariance?
\item The modular invariance is important in the number theory. What problems of number theory are connected through the modular transformations with the representation theory?
\item I need to write about the Hecke algebras
\item How are Hecke algebras connected to the modular forms?
\item Are there any kind of modular invariance analogues for the finite-dimensional Lie algebras? There exists the lattice.
\item What become of modular invariance during the quantization? Are there any properties of quantum integrable systems that follow from the modular invariance?
\end{itemize}
